Thesis outline

### 2017-06-08 KellyDouglass/Thesis/Thesis_sections/smallScaleEnvironment.tex
Updated smallScaleEnvironment chapter with Steve's comments

% Large-scale environment influence
Large galaxy redshift surveys have shown that the large-scale structure of the 
distribution of galaxies in the universe is similar to that of a 
three-dimensional cosmic web \citep{Bond96}, where galaxy clusters are connected 
by thin filaments of galaxies and are separated by voids (large, underdense 
regions that occupy close to 60\% of space).  Over the last few decades, the 
Sloan Digital Sky Survey \citep{Abazajian09,Ahn12} have propelled the research 
into the influence of the large-scale environment on the formation and evolution 
of galaxies.

Because gravitational clustering within a void proceeds as if in a very 
low-density universe, cosmic voids are an important environment for studying 
galaxy formation \citep[see][for a review]{vandeWeygaert11}.  The $\Lambda$CDM 
cosmology predicts that galaxies which formed in voids to have lower masses and 
be retarded in their star formation when compared to those in more dense 
environments \citep{Gottlober03,Goldberg05,Cen11}.  The effects of the void 
environment should be most obvious in dwarf galaxies due to their minimal 
gravitational potential.  As a result, they are more sensitive to astrophysical 
effects such as cosmological reionization, internal feedback from supernova and 
photoheating from star formation, small-scale details of dark matter halo 
assembly, and the properties of dark matter.

@@

% Small-scale environment - what has been done so far?
In conjunction with the large-scale environment, the small-scale environment 
($\sim 1$ \hMpc) has also been found to influence a galaxy's evolution.  There 
exits a well-established morphology-density relation \cite{Dressler80}, where 
the fraction of late-type galaxies is inversely proportional to the local 
density.  \cite{Ellison09} determined that a galaxy's local environment 
influences a galaxy's evolution more than its large-scale environment.  
Likewise, \cite{Rupke08} concluded that interacting galaxies have suppressed 
metallicities, because interactions induce flows of hydrogen.  \cite{Park09} 
found that, for galaxies with $M_r < -19$, galaxy interactions out to the virial 
radius of the nearest neighbor influence the evolution of the target galaxy.  
They determine that the large-scale environment has a minimal effect on the 
evolution of a galaxy once the luminosity and morphology are taken into account.

We want to understand if the small-scale environment is more influential in a 
galaxy's evolution than its large-scale environment.  Additionally, within the 
small-scale environment, are tidal influences (due to the nearest neighbor 
galaxy) or gravitational potentials (from the nearest galaxy group) more 
influential in a galaxy's evolution?  A group consists of multiple galaxies that 
share the same dark matter halo.  A group member will gravitationally interact 
with all other group members, not just its nearest neighbor.  And a galaxy near 
a group, but not within it, may also experience strong tidal effects from the 
group similar to (but stronger than) those from a neighboring galaxy.

@@

We employ two different spatial metrics to determine the best way to locate a 
galaxy's nearest neighbor.  Within a relative velocity 
$v_{rel} < 300\text{ km/s}$, we use the physically closest (in units of \hMpc) 
and the neighbor with the smallest fraction of its virial radius; both these 
methods are described below in further detail.  The two methods result in 
different neighboring galaxies for 40\% of our dwarf galaxy sample.

@@

Rather than using a measure of the virial radius for the group analysis, we 
scale the distance as a fraction of the group's rms radius.

@@

The relationships between various physical parameters and the distance to the 
nearest neighbor or group for our sample of star-forming dwarf galaxies are 
shown in Figs. \ref{fig:ur}--\ref{fig:NO}.  In each of the four plots within 
each figure, we bin the galaxies in distance-space and show the average 
parameter value in each bin, to see any overall trends in the data.  We also fit 
a linear line to the data; the output of these fits are in Table \ref{tab:fits}.  
Each of these four plots probes this distance relationship from a different 
angle.  The top left plots in Figs. \ref{fig:ur}--\ref{fig:NO} compare the 
target dwarf galaxy's parameter to the distance to its nearest neighbor, as 
defined by the the closest galaxy on the sky in \hMpc with a velocity difference 
less than 300 km/s.  The bottom left plots compare the target dwarf galaxy's 
parameter to the distance to its nearest neighbor, as defined by the closest 
galaxy on the sky in units of the virial radius of the neighbor galaxy with a 
velocity difference less than 300 km/s.  The nearest galaxy is not necessarily 
the same for these two distance measurements, as explained in Sec. 
\ref{sec:Theory_dist}.

We repeat this same analysis on the nearest groups to the target dwarf galaxy.  
The top right plots in Figs. \ref{fig:ur}--\ref{fig:NO} compare the target dwarf 
galaxy's parameter to the distance to the center of the nearest group, as 
defined as the closest group on the sky in \hMpc with a velocity difference less 
than 300 km/s.  The bottom right plots in Figs. \ref{fig:ur}--\ref{fig:NO} 
compare the target dwarf galaxy's parameter to the distance to the nearest group 
as defined by the closest group on the sky in units of the group's rms radius 
with a velocity difference of less than 300 km/s.  Groups are very rare in the 
void environment (by nature of the void environment), so the uncertainty in the 
binned values shown in these plots is much larger than in the galaxy neighbor 
plots.

@@

Because of the known morphology-density relation, we excepted to find that a 
dwarf galaxy's color became more blue as the distance to its nearest neighbor 
increased.  Fig. \ref{fig:ur} shows very little relationship between the 
distance and color, except at the smallest distance bin.  The linear fits 
quantify this observation --- the slopes are on the order of 10$^{-3}$.  
However, within a distance of 0.05 \hMpc or 0.05$r_{vir}$, dwarf galaxies tend 
to be bluer than at further distances from their nearest neighbor.  At distances 
less than 0.1 \hMpc from the center of the nearest groups, the dwarf galaxies 
are redder than average.  However, there does not appear to be any relationship 
between a dwarf galaxy's color and its distance to the center of the nearest 
group in units of the group's rms radius.  It has been well-established that 
void galaxies tend to be bluer than galaxies in denser environments 
\citep{Grogin99,Rojas04,Patiri06,vonBendaBeckmann08,Hoyle12}; this shift is 
apparent in Fig. \ref{fig:ur}, where the void dwarf galaxies are slightly bluer 
than the wall dwarf galaxies.

@@

We test two of our input parameters to understand how sensitive our results are 
to any initial conditions.  The first parameter we discuss is the sensitivity of 
our results on the maximum peculiar velocity to define a match.  Throughout our 
analysis, we use 300 km/s as the maximum velocity separation allowed between the 
target galaxy and its nearest neighbor or group.  We look at how this affects our 
results by repeating the analysis with maximum velocities of 150 km/s and 600 
km/s.  The results of this comparison on the color of the galaxies can be seen in 
Fig. \ref{fig:ur_vpeculiar}.  The nearest neighbors in the left-hand panel are 
restricted to a maximum peculiar velocity of 150 km/s, while those in the right 
panel are restricted to 600 km/s.  When compared to each other and to the 
left-hand panel of Fig. \ref{fig:ur}, it is clear that our choice of maximum 
peculiar velocity has no affect on the results of the study.

@@

We note the existence of a gap in the distribution of galaxies around a distance 
of 0.05 \hMpc or 0.5$r_{vir}$ from the nearest neighbor that is not present in 
Fig. \ref{fig:ur}.  When we incorporate the redshift into the distance 
calculations, the resulting distance between galaxies which have small sky 
separations but larger redshift separations is much larger than those with 
larger sky separations and little redshift separation.  As a result, galaxies 
which were originally close to the color axis in Fig. \ref{fig:ur} will be 
moved to much larger distances, while the locations of those which were 
originally further from the color axis will change much less.  The dwarf 
galaxies which remain close to the color axis in Fig. \ref{fig:ur_xyz} must, 
therefore, have small sky separation and almost no difference in their peculiar 
velocities.  We conclude that these represent merging systems, which visual 
inspection will help to confirm.

@@

We see almost no relationship between a dwarf galaxy's color, sSFR, or gas-phase 
chemical abundance and its distance to the nearest galaxy or group, implying 
that the small-scale environment does not significantly influence a dwarf 
galaxy's evolution.  This is in contrast to the large-scale environment, which 
we have seen to influence the formation and evolution of dwarf galaxies.

Only those galaxies within 0.05 \hMpc and 0.05$r_{vir}$ of a nearest neighbor, 
or those within 0.1 \hMpc of the nearest group appear to deviate from the 
average galaxy values.  The galaxies within this proximity of their nearest 
neighbor tend to be bluer, have a higher sSFR, and have higher oxygen and 
nitrogen abundances.  Based on the shift in the distribution of galaxies in Fig. 
\ref{fig:ur_xyz}, these galaxies might be merging or strongly interacting with 
their nearest neighbor.  This would then provide evidence that galaxy 
interactions result in a burst of star formation, which would increase the 
gas-phase chemical abundances of the dwarf galaxies.  Because merging galaxies 
share the same dark matter halo, it appears that the sharing of a dark matter 
halo has more influence on the evolution of a galaxy than the distance to its 
nearest neighbor.

Dwarf galaxies within 0.1 \hMpc of the center of the nearest group are redder, 
have lower oxygen abundances, and have higher N/O ratios than average.  Being so 
close to the center of a group prevents more recent episodes of star formation.  
Due to their proximity to the group center, it is also likely that these dwarf 
galaxies are not able to retain as much of their heavier elements as for a more 
isolated galaxy, reducing their gas-phase oxygen abundance (and therefore 
increasing their N/O ratio).

>>>>

% Large-scale environment influence
Large galaxy redshift surveys have shown that the large-scale structure of the 
distribution of galaxies in the universe is similar to that of a 
three-dimensional cosmic web \citep{Bond96}, where galaxy clusters are connected 
by thin filaments of galaxies and are separated by voids (large, underdense 
regions that occupy close to 60\% of space).  Over the last few decades, the 
Sloan Digital Sky Survey \citep{Abazajian09,Ahn12} has accelerated the research 
into the influence of the large-scale environment on the formation and evolution 
of galaxies.

Because gravitational clustering within a void proceeds as if in a very 
low-density universe, cosmic voids are an important environment for studying 
galaxy formation \citep[see][for a review]{vandeWeygaert11}.  The $\Lambda$CDM 
cosmology predicts that galaxies formed in voids have lower masses and are 
retarded in their star formation when compared to those in more dense 
environments \citep{Gottlober03,Goldberg05,Cen11}.  The effects of the void 
environment should be most obvious in dwarf galaxies due to their minimal 
gravitational potential.  As a result, they are more sensitive to astrophysical 
effects such as cosmological reionization, internal feedback from supernovae and 
photoheating from star formation, small-scale details of dark matter halo 
assembly, and the properties of dark matter.

@@

% Small-scale environment - what has been done so far?
In conjunction with the large-scale environment, the small-scale environment 
($\sim 1$ \hMpc) has also been found to influence a galaxy's evolution.  A 
well-established morphology-density relation exists \citep{Dressler80}, where 
the fraction of late-type galaxies is inversely proportional to the local 
density.  \cite{Ellison09} determined that a galaxy's local environment 
influences a galaxy's evolution more than its large-scale environment.  
Likewise, \cite{Rupke08} concluded that interacting galaxies have suppressed 
metallicities, because interactions induce flows of hydrogen.  \cite{Park09} 
found that, for galaxies with $M_r < -19$, galaxy interactions out to the virial 
radius of the nearest neighbor influence the evolution of the target galaxy.  
They determine that the large-scale environment has a minimal effect on the 
evolution of a galaxy once the luminosity and morphology are taken into account.

We want to understand if the small-scale environment is more influential in a 
galaxy's evolution than its large-scale environment.  One important question 
within the small-scale environment asks if tidal influences due to the nearest 
neighbor galaxy or gravitational potentials from the nearest galaxy group are 
more influential in a galaxy's evolution.  A group consists of multiple galaxies 
that share the same dark matter halo.  A group member will interact 
gravitationally with all other group members, not just its nearest neighbor.  
And a galaxy near a group, but not within it, may also experience strong tidal 
effects from the group similar to and stronger than those from a neighboring 
galaxy.

@@

We employ two different spatial metrics to determine the best way to locate a 
galaxy's nearest neighbor.  Within a relative velocity 
$v_{rel} < 300\text{ km/s}$, we use either the physically closest galaxy in 
units of \hMpc or the galaxy within the smallest fraction of its virial radius; 
both these methods are described below in further detail.  The two methods 
result in different neighboring galaxies for about 40\% of our dwarf galaxy 
sample.

@@

Rather than using a measure of the virial radius for the group analysis, we 
scale the distance as a fraction of the group's root mean square (rms) radius.

@@

The relationships between various physical parameters and the distance to the 
nearest neighbor or group for our sample of star-forming dwarf galaxies are 
shown in Figs. \ref{fig:ur}--\ref{fig:NO}.  In each of the four plots within 
each figure, we bin the galaxies in distance-space and show the average 
parameter value in each bin, to see any overall trends in the data.  We also fit 
a linear line to the data; the output of these fits are in Table \ref{tab:fits}.  
Each of these four plots probes this distance relationship from a different 
angle.  The top left plots in Figs. \ref{fig:ur}--\ref{fig:NO} relate the target 
dwarf galaxy's parameter to the distance to its nearest neighbor, as defined by 
the the closest galaxy on the sky in \hMpc with a velocity difference less than 
300 km/s.  The bottom left plots relate the target dwarf galaxy's parameter to 
the distance to its nearest neighbor, as defined by the closest galaxy on the 
sky in units of the virial radius of the neighbor galaxy with a velocity 
difference less than 300 km/s.  The nearest galaxy is not necessarily the same 
for these two distance measurements, as explained in Sec. \ref{sec:Theory_dist}.

We repeat this same analysis on the nearest groups to the target dwarf galaxy.  
The top right plots in Figs. \ref{fig:ur}--\ref{fig:NO} relate the target dwarf 
galaxy's parameter to the distance to the center of the nearest group, as 
defined as the closest group in the sky in \hMpc with a velocity difference less 
than 300 km/s.  The bottom right plots in Figs. \ref{fig:ur}--\ref{fig:NO} 
relate the target dwarf galaxy's parameter to the distance to the nearest group 
as defined by the closest group on the sky in units of the group's rms radius 
with a velocity difference of less than 300 km/s.  Groups are very rare in the 
void environment (by nature of the void environment), so the uncertainty in the 
binned values shown in these plots is much larger than in the galaxy neighbor 
plots.

@@

Because of the known morphology-density relation, we expect to find that a dwarf 
galaxy's color became more blue as the distance to its nearest neighbor 
increased.  Fig. \ref{fig:ur} shows very little relationship between the 
distance and color, except at the smallest distance bin.  The linear fits 
quantify this observation --- the slopes are on the order of 10$^{-3}$.  
However, within a distance of 0.05 \hMpc or 0.05$r_{vir}$, dwarf galaxies tend 
to be bluer than at further distances from their nearest neighbor.  At distances 
less than 0.1 \hMpc from the center of the nearest groups, the dwarf galaxies 
are redder than average.  However, there does not appear to be any relationship 
between a dwarf galaxy's color and its distance to the center of the nearest 
group in units of the group's rms radius.  It has been well-established that 
void galaxies tend to be bluer than galaxies in denser environments 
\citep{Grogin99,Rojas04,Patiri06,vonBendaBeckmann08,Hoyle12}; this shift is 
apparent in Fig. \ref{fig:ur}, where the void dwarf galaxies are slightly bluer 
than the wall dwarf galaxies.

@@

We test two components of our nearest neighbor criteria to understand how 
sensitive our results are to any initial conditions.  The first parameter we 
discuss is the sensitivity of our results on the maximum peculiar velocity to 
define a match.  Throughout our analysis, we use 300 km/s as the maximum 
velocity separation allowed between the target galaxy and its nearest neighbor 
or group.  We look at how this affects our results by repeating the analysis 
with maximum velocities of 150 km/s and 600 km/s.  The results of this 
comparison on the color of the galaxies can be seen in Fig. 
\ref{fig:ur_vpeculiar}.  The nearest neighbors in the left-hand panel are 
restricted to a maximum peculiar velocity of 150 km/s, while those in the right 
panel are restricted to 600 km/s.  When compared to each other and to the 
left-hand panel of Fig. \ref{fig:ur}, it is clear that our choice of maximum 
peculiar velocity has no affect on the results of the study.

@@

We note the existence of a gap in the distribution of galaxies around a distance 
of 0.05 \hMpc or 0.5$r_{vir}$ from the nearest neighbor that is not present in 
Fig. \ref{fig:ur}.  When we incorporate the redshift into the distance 
calculations, the resulting distance between galaxies which have small sky 
separations but larger redshift separations is much larger than those with 
larger sky separations and little redshift separation.  As a result, galaxies 
which were originally close to the color axis in Fig. \ref{fig:ur} will be 
moved to much larger distances, while the locations of those which were 
originally further from the color axis will change much less.  The dwarf 
galaxies which remain close to the color axis in Fig. \ref{fig:ur_xyz} must, 
therefore, have small sky separation and almost no difference in their peculiar 
velocities.  We surmise that these represent merging systems, which future 
visual inspection will help to confirm.

@@

We see no relationship between a dwarf galaxy's color, sSFR, or gas-phase 
chemical abundances and its distance to the nearest galaxy or group beyond the 
target galaxy's immediate vicinity, implying that the small-scale environment 
does not significantly influence a dwarf galaxy's evolution.  This is in 
contrast to the large-scale environment, which we have seen to influence the 
formation and evolution of dwarf galaxies.

Only those galaxies within 0.05 \hMpc and 0.05$r_{vir}$ of a nearest neighbor, 
or those within 0.1 \hMpc of the nearest group appear to deviate from the 
average galaxy values.  The target galaxies within this proximity of their 
nearest neighbor tend to be bluer, have a higher sSFR, and have higher oxygen 
and nitrogen abundances.  Based on the shift in the distribution of galaxies 
seen in Fig. \ref{fig:ur_xyz}, these galaxies might be merging or strongly 
interacting with their nearest neighbor.  If so, this provides evidence that 
galaxy interactions result in a burst of star formation that increases the 
gas-phase chemical abundances of the dwarf galaxies.  Because merging galaxies 
share the same dark matter halo, it appears that the sharing of a dark matter 
halo has more influence on the evolution of a galaxy than the distance to its 
nearest neighbor.

In contrast, dwarf galaxies within 0.1 \hMpc of the center of the nearest group 
are redder, have lower oxygen abundances (O/H), and have higher N/O ratios than 
average.  Being so close to the center of a group seems to prevent more recent 
episodes of star formation.  Due to their proximity to the group center, it is 
also likely that these dwarf galaxies are not able to retain as much of their 
heavy elements as a more isolated galaxy, thereby reducing their gas-phase 
oxygen abundance (and increasing their N/O ratio).
